### 2015-12-19 Rmd/lit-review-analyses.Rmd
Commit of full datasets, some moving around of files, first package function added (kdensity.R).

---
title: "Supporting analysis of paper metadata"
author: "Lyndon Estes"
date: "February 24, 2015"
output: html_document
---

From 10.5194/bg-11-463-2014. Figuring out parameters from MODIS data used in study. 
```{r}
library(sp)
library(rgeos)
library(raster)
library(geosphere)
library(rworldmap)
data(countriesLow)

r <- raster(extent(-70, 30, -55, -25))
res(r) <- 0.00833
areas <- area(r)
plot(areas)
areas_m <- mask(areas, countriesLow, inverse = TRUE)
plot(areas_m)
plot(countriesLow, add = TRUE)
cellStats(areas_m, sum) * 100  # 23795817
cellStats(areas_m, mean)
1000^2 * 0.64005
freq(areas_m)
```