### 2015-12-19 Rmd/select-titles.Rmd

---
title: "Select titles"
author: "Lyndon Estes"
date: "December 19, 2015"
output: 
  html_document:
    toc: yes
---

# Select papers
Organization and random selection of titles from literature review.

```{r, eval=FALSE}
library(lmisc)
p_root <- full_path(proj_root(proj.dir = "lit_review"), "lit_review")
p_data <- full_path(p_root, "data")

# Read in downloaded citations and merge
batch1 <- read.csv(full_path(p_data, "top30_from1982_set1.csv"))  
batch2 <- read.csv(full_path(p_data, "top30_from1982_set2.csv"))
articles <- rbind(batch1, batch2)

# Shuffle full list 
set.seed(234)
draw <- sample(1:nrow(articles), size = nrow(articles), replace = FALSE)
article.shuffle <- articles[draw, ]

# 10-year list\set.seed(234)
articles.2004 <- articles[articles$Year >=2004, ]

set.seed(234)
draw <- sample(1:nrow(articles.2004), size = nrow(articles.2004), 
               replace = FALSE)
article.2004.shuffle <- articles.2004[draw, ]

# draw subsets for each reviewer - note: as long as drawsize and random seed are the same, new reviewers can 
# be appended to the group without affecting the draw for existing reviewers
reviewers <- c("Chang", "Treuer", "Socolar", "Daskin", "Elsen", "Estes", 
               "Choi", "Margulies", "JChang", "Ahmed")
nauth <- length(reviewers)
#drawsize <- (round(nrow(articles.2004) * 0.025 / nauth) * nauth) / nauth  # number divisible by 
drawsize <- 400
set.seed(432)
sset <- sample(1:nrow(article.2004.shuffle), size = drawsize * nauth, 
               replace = FALSE)
inds <- seq(1, length(sset) + 1, by = length(sset) / nauth)
ind.mat <- cbind(inds[-length(inds)], 
                 c(inds[-c(1, length(inds))] - 1, length(sset)))

# New calibration set randomly selected (note, run on earlier version of random draw taken before Elsen, 
# so to replicate use commented out drawsize and remove Elsen from reviewers before re-running)
# article.2004.shuffle.red <- article.2004.shuffle[-sset, ]
# set.seed(124)
# cal.set <- article.2004.shuffle.red[sample(1:nrow(article.2004.shuffle.red), size = 20, replace = FALSE), ]
# # which(rownames(cal.set) %in% rownames(article.2004.shuffle[sset, ]))  # 0
# write.csv(cal.set, file = "new_calibration_set.csv")

# Then draw sets for each reviewer
fnames <- full_path(p_data, paste(reviewers, "2004_draw3.csv", sep = "_"))
sel.set <- vector("list", length = length(fnames))
names(sel.set) <- reviewers
for(i in 1:nrow(ind.mat)) {
  sel.set[[i]] <- article.2004.shuffle[sset[ind.mat[i, 1]:ind.mat[i, 2]], ]
  write.csv(sel.set[[i]], file = fnames[i])
}

# check that rownames aren't duplicated
sel.set[[1]][1:10, ]
for(i in 1:(length(sel.set) - 1)) {
   print(any(sel.set[[i]]$Title == sel.set[[i + 1]]$Title))
}

fn_old <- dir(p_data, 
              pattern = paste0("^", 
                               paste(reviewers[1:6], "draw2.csv", 
                                     sep = ".*.", collapse = "|")))
fn_old <- fn_old[sapply(reviewers[1:6], function(x) grep(x, fn_old))]
fn_new <- dir(p_data, 
              pattern = paste0("^", 
                               paste(reviewers[1:6], "draw3.csv", 
                                     sep = ".*.", collapse = "|")))
fn_new <- fn_new[sapply(reviewers[1:6], function(x) grep(x, fn_new))]

# comparing draw2 and draw3
for(i in 1:length(fn_old)) {
  old <- read.csv(full_path(p_data, fn_old[i]))
  neww <- read.csv(full_path(p_data, fn_new[i]))
  print(which(!old$Title %in% neww$Title))
}  # only last record of old set is not in new set (b/c bad indexing)

# comparing draw2 titles in sequence
for(i in 1:(length(fn_old) - 1)) {
  old <- read.csv(full_path(p_data, fn_old[i]))
  oldp1 <- read.csv(full_path(p_data, fn_old[i + 1]))
  print(paste("common title", which(old$Title %in% oldp1$Title)))
  print(which(old$X %in% oldp1$X))
}  # overlap between last of preceding and 1st of next (and if no title)

fnames_newall <- dir(p_data, pattern = "draw3.csv")
for(i in 1:(length(fn_new) - 1)) {
  neww <- read.csv(full_path(p_data, fn_new[i]))
  newp1 <- read.csv(full_path(p_data, fn_new[i + 1]))
  print(which(neww$X %in% newp1$X))
  print(which(neww$Title %in% newp1$Title))
}  # no duplicate names in new file sets (unless title is untitled)

```