### 2016-04-02 Rmd/calibration.Rmd
Bulk commit of various changes, including library functions added. Edits to space-time.Rmd include near-final version of figure.

---
title: "calibration"
author: "Lyndon Estes"
date: "January 8, 2016"
output: 
  html_document:
    toc: yes
    fig_caption: true
---

# Calibration data
```{r, message=FALSE, warning = FALSE}
library(ecoscales)
library(readxl)
library(Agreement)
library(irr)

p_root <- set_base_path("ecoscales")
p_dat <- fp(p_root, "external/data/calibration")

# Read in and clean-up calibration data
# cal <- data.table(read_excel(fp(p_dat, "merged_calibration.xlsx")))
cal <- data.table(read_excel(fp(p_dat, "merged_calibrationff.xlsx")))
cal[, DOI := gsub("DOI:|doi:|DOI: |doi: |DOI ", "", DOI)]
cal[, DOI := gsub("\\s", "", DOI)]  # remove DOI and whitespace
dropnm <-  c("V30", "V31", "journal", "exclusion reason", "sensitivity")
cal[, c(dropnm) := NULL]

# check data types
for(j in names(cal)[8:24]) set(cal, j = j, value = as.numeric(cal[[j]]))

# use NA in n_sites as indicator that study was excluded
cal[, inout := ifelse(is.na(n_sites), 0, 1)]  
# write.csv(cal, file = fp(p_dat, "merged_calibrationf.csv"))

# external fixes and recodes, bring back in
cal <- data.table(read_excel(fp(p_dat, "merged_calibrationf.xlsx")))
cal[, notes := NULL]
for(j in names(cal)[c(4:5, 8:24)]) set(cal, j = j, value = as.numeric(cal[[j]]))
# length(which(is.na(cal$feature)))

```

## Agreement on studies included/excluded
```{r}
# ndois <- cal[, unique(DOI)]  # check on number of journals

# reduce within-observer DOI duplicates 
dois <- cal[, unique(DOI), by = .(observer, inout)]

# recast dataset to obsevers in columns, inout as value, DOI as case
dois_c1 <- dcast(dois, V1 ~ observer, value.var = "inout")

# Fleiss's Kappa
kappam.fleiss(dois_c1[, 2:7, with = FALSE])
```

## Inter-rater reliability using number of records per paper
```{r}
doisum <- dcast(cal, DOI ~ observer, value.var = "inout", fun = sum)
# kappam.fleiss(doisum[, 2:7, with = FALSE])
icc(ratings = doisum[, 2:7, with = FALSE], model = "twoway")

```

```{r, warning = FALSE, fig.cap="Figure 1: Number of records extracted by each rater (observer) for each of the 20 papers in the calibration set."}
doimat <- t(doisum[, -1, with = FALSE])
cols <- RColorBrewer::brewer.pal(n = 6, "Paired")#[c(3:1, 4:6)]
cxs <- seq(3, 0.8, length.out = 6)
pchs <- c(rep(20, 5), 4)
par(mar = c(4, 4, 0.1, 0.1))
plot(c(1, 20), c(0, 10), pch = "", xlab = "Paper reviewed", 
     ylab = "N extractable records", mgp = c(1.45, 0.7, -0.2))
polygon(x = c(0, 21, 21, 0, 0), y = c(-1, -1, 11, 11, 0), col = "grey70")
# plot(doimat[1, ], pch = 20, xlab = "Paper reviewed", 
#      ylab = "N extractable records", cex = cxs[1], col = cols[1])
for(i in 1:6) points(doimat[i, ], col = cols[i], pch = pchs[i], cex = cxs[i])
legend(x = "topleft", pch = pchs, pt.cex = cxs, col = cols, 
       legend = paste0("Rater", 1:6), bty = "n")

```

## Variance in extracted statistics
```{r}
varnms <- c("plot_res", "n_sites", "sampled_area", "t_btwn_samp", 
            "study_duration")
mus <- cal[inout == 1, lapply(.SD, mean, na.rm = TRUE), by = .(DOI, observer),             .SDcols = varnms]
vars <- mus[, lapply(.SD, sd, na.rm = TRUE), by = DOI, .SDcols = varnms]
musmu <- mus[, lapply(.SD, mean, na.rm = TRUE), by = DOI, .SDcols = varnms]

par(mfrow = c(3, 2), mar = c(3, 3, 2, 2), mgp = c(1.5, 0.5, 0))
for(i in varnms) {
  mu <- musmu[, get(i)]
  sdev <- vars[, get(i)]
#   print(paste(i, ":", sdev))
  print(paste(i, ":", round(mu, 2), ":", round(sdev, 2), ":", 
              round(sdev / mu, 2)))
  # plot(mu, sdev / mu, xlab = "mean", ylab = "CV", main = i)
  plot(mu, sdev, xlab = "mean", ylab = "sd", main = i)
}

vars[, lapply(.SD, mean, na.rm = TRUE), .SDcols = varnms]
colMeans(sapply(varnms, function(i) {
  mu <- musmu[, get(i)]
  sdev <- vars[, get(i)]
  sdev / mu
}), na.rm = TRUE)

```

```{r}
par(mfcol = c(10, 4), mar = c(2, 2, 2, 2), mgp = c(1.5, 0.5, 0))

mus[, plot(plot_res), by = DOI]

```

Tim needs to fix paleo studies in calibration (and main set) giving too much credit to effective area.

Others need to fix their study duration figures. Paul on paleo one in calibration, Choi, Chang, etc in main set.

For N-sites, count sites only once if there are repeated measures taken at them, e.g. Jonathan in 10.1111/j.1466-8238.2009.00454.x for bird counts