### 2018-03-31 README.Rmd
Edited README

## Overview

This package contains the data and code for the following paper:

Estes, L., Elsen, P., Treuer, T., Ahmed, L., Caylor, K., Chang, J., Choi, J.J., Ellis, E. (2018). The spatial and temporal domains of modern ecology. __Nature Ecology and Evolution__, DOI:10.1038/s41559-018-0524-4

The published version of the paper can be accessed [here]( http://dx.doi.org/10.1038/s41559-018-0524-4).

## Installation

The package can be installed and contents explored as follows:

```{r, eval = FALSE}
devtools::install_github("agroimpacts/ecoscales", build_vignettes = TRUE)
browseVignettes("ecoscales")
```